---
title: "Database project assignment"
author: "Lucas Antelo"
date: "May 10, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Modeling

###Executive Summary
--> TODO

###Getting the Data and Exploration
For this assignment we will use the dataset american-election-tweet.xlsx downloaded form <http://kvv.imp.fu-berlin.de> under the subsection ressources from the database course. In this section we try to explore the data set.

2. Aufgabe: Explorative Datenanalyse
  1.  Laden Sie sich den Datensatz "american-election-tweets.csv" oder "american-
      election-tweets.xlsx" aus dem KVV herunter.
  2.  Schauen Sie sich den Datensatz an und versuchen Sie die Zusammenhange
      im Datensatz zu verstehen. Der Datensatz enthalt mehr Daten als fur
      das Projektziel relevant sind.
  3.  Beschreiben Sie den Datensatz.

```{r overview, warning=FALSE}
#the data for the assignemnt is in the subfolder /data
#file <- "data/american-election-tweet.csv"
#data <- read.csv(file) fails as there are more columns than column names
library(readxl)
data <- read_excel("data/american-election-tweets.xlsx")
#what attributes are available?
colnames(data)
#dimension
dim(data)
#structur of data
str(data)

#summary of the data
summary(data)
```
The data set consists of 6126 entries and 11 attributs. We will look deeper into each of these attributes and moreover explore any correlations, distributions and other interesting facts, that we can analyze.

## Anaylzing the the attributes

We have eleven attributes consisting with the following values and characteristics: 

```{r attributes, eval=FALSE }
unique(data$handle) # "HillaryClinton"  "realDonaldTrump"
unique(data$text) # many different texts -> further analysis required
unique(data$is_retweet) # true false
unique(data$original_author) # 268 different authors -> further analysis required
range(data$time) # "2016-01-05T03:36:53" "2016-09-28T00:22:34"
unique(data$in_reply_to_screen_name) 
# NA "HillaryClinton" "Lesdoggg" "elizabethforma" "MaryKayHenry" "montgomeriefdn" "JLo" "POTUS" "tomcolicchio"  
unique(data$is_quote_status) # true false
range(data$retweet_count) # 123 490180
range(data$favorite_count) # 274 660384
unique(data$source_url) 
# "https://studio.twitter.com"                   "http://twitter.com"                          
# "https://about.twitter.com/products/tweetdeck" "http://twitter.com/download/iphone"          
# "http://twitter.com/download/android"          "http://twitter.com/#!/download/ipad"         
# "http://instagram.com"                         "https://mobile.twitter.com"                  
 
unique(data$truncated) # true false
```
The data set consist of entries that start on January 5 2016 and ends on September 28 2016. //domain

We see that the attributes
  handle, is_retweet, is_quote_status and truncated       //domain of each one
consist of binary values - in other words there only are two possibilities

The attribute text is much more verbose and will need further analysis. //domain

There are 268 unique orignal_author entries. 8 unique in_reply_to_scren_name. //domain 

The range of retweet_count is 123 to 490180.  //not a domain but range is better
The range of favorite_count is 274 to 660384. //not a domain but range is better

There are 8 unique source_url entries.

## 2. Data Import

## 3. Data Mining and Visualization
